Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}

### 2023-12-02 draft/draft.tex
draft

\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference and the University of Reading. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
\author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
\date{2023-12-01}

@@

To identify the effects of air pollution on agricultural productivity, I identify areas of high pollution emission using the location of coal plants. While coal plants may add emissions, I show that they are located in areas with relatively higher levels of pollution, in general. Using these locations as point sources of pollution, I use plausibly exogenous changes in wind direction -- in a similar spirit to Deryugina et al. (2019) -- to measure the effects of pollution on key agricultural outcomes. I first plot the location of coal plants in India by year. Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 2000s to the 2010s, with differences across outcomes depending on data availability. I then calculate the direction from each coal plant to all villages within 30km. I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution later in the paper.

@@

To measure exposure to pollution from point sources around coal plants, I first locate all village centroids located within 30km from a coal plant in a given year. I choose 30km due to previous research on the effects of (air) pollution (Aragón and Rud 2016; Bencko and Symon 1977; Li and Gibson 2014). I then calculate the direction from these point sources to all village centroids within that 30km radius. To define exposure, I pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids. The wind data includes both the x and y components of wind speed, which I can use to calculate angles from point sources to village centroids.} and then temporally aggregate this daily data depending on the temporal definition of the corresponding outcome. For example, agricultural productivity is defined across five months (i.e.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.

@@

While I use PM 2.5 as the estimated level of pollution, it is important to clarify that I cannot explicitly say that PM 2.5, per se, is responsible for any effects. Pollution includes a variety of different components, including PM 2.5, PM 10, sulfur dioxide, and nitrogen dioxide, to name but a few. While I use PM 2.5 as a proxy for pollution, it is possible that other components are responsible for the effects. Additionally, the PM 2.5 data is estimated using satellite data and a machine learning algorithm. While this is a common approach in the literature ( see Hammer et al. (2020) for more details), it is important to note that the estimates are not direct measurements of pollution. However, a valid instrumental variable can help for any measurement error present in the data (Glewwe 2012; Gillen, Snowberg, and Yariv 2019).

@@

What are the characteristics of areas near the construction of new coal plants? Table \ref{tab:plantresultstable} presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants. The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level. The first column presents results from regressing a dummy for whether there is a coal plant within 30km in 1991 on several census variables. Villages with higher population densities (higher log population and lower log area) are more likely to be near a coal plant at the time. Urban population proportion is not a significant predictor after conditioning on population and area.

@@

The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there \emph{will be} a coal plant constructed within 30km before 2001. This column restricts the sample to only villages that do not already have a coal plant within 30km in 1991. The pattern of significance remains the same in column two.

The last two columns are particularly relevant to the current results. They present similar results to those in the first two columns, but use the 2001 census data and a dummy for the presence of a coal plant in 2001 and 2011 in columns three and four, respectively. Importantly, I include agricultural productivity in the monsoon season of 2002 -- the earliest year available -- as well as average pollution during the monsoon season for 1998 through 2000 as additional predictors. In 2001, agricultural productivity is a strong predictor for the presence of a coal plant. Moreover, it is also a significant predictor of the construction of a coal plant between 2001 and 2011 (column four). Given that the coefficient is positive, this indicates that coal plants are more likely to be built in areas with higher agricultural productivity. In other words, any selection in the overall location of coal plants would bias against finding negative effects of pollution on agricultural productivity. However, note that this is not the same as looking at wind direction, though comparing the OLS and IV results below leads to the same conclusion regarding selection.

Pollution is also a significant predictor for the presence of a coal plant in 2001 as well as for whether a coal plant will be constructed before 2011. While the coefficient is relatively small, it is actually quite large given that only three percent of villages in column four will get a new plant in the next decade.

@@

It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution. Consider the data used for the regressions presented in Table \ref{tab:yield3ivmainfirststage} of the appendix. The outcome variable is particulate matter -- specifically, \(\mathrm{PM 2.5}\), which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from Hammer et al. (2020). \(\mathrm{PM 2.5}\) is one of the harmful byproducts from many industrial processes, including power generation. Particulate matter is a major component of the pollution emitted by coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, but I aggregate variables to the season variable since that is the level of analysis in the main results.

Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \ref{tab:yield3ivmainfirststage} are at the village level. As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are month fixed effects. Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level. I add more and more flexible specifications for weather, with the most flexible specification in the fourth column. Note that the relationship between wind direction and particulate matter is very similar across all four columns, suggesting that the inclusion of weather does not change the relationship between wind direction and pollution. This is important because it suggests that the use of wind direction as a proxy for pollution exposure is not biased by weather patterns.

@@

Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity. This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards: although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of pollution sources in the study areas. One possible way to recover the actual effect of pollution on agricultural productivity is with an instrument. If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.

@@

\caption{\label{tab:yieldtableleads}Agricultural productivity and pollution leads}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
\toprule
  & (1) & (2)\\
\midrule
particulate matter (one-year lead) & -0.033 & \\
 & (0.067) & \\
particulate matter (two-year lead) &  & -0.070\\
 &  & (0.052)\\
weather (expanded) & No & No\\
\textbf{fixed effects:} & \textbf{} & \textbf{}\\
village & Yes & Yes\\
year & Yes & Yes\\
\midrule
F & 592 & 783\\
observations & 1,161,265 & 1,055,562\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\hypertarget{crops-and-labor-intensity}{%
\subsection{Crops and labor intensity}\label{crops-and-labor-intensity}}

@@

\caption{\label{tab:yieldrice}Agricultural productivity by most common crop}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
  & (1) & (2)\\
\midrule
Particulate matter (log PM 2.5) & -0.287** & -0.394***\\
 & (0.133) & (0.134)\\
PM 2.5 times rice & -0.338*** & -0.205***\\
 & (0.061) & (0.060)\\
Weather (expanded) & Yes & No\\
Weather (expanded, bins) & No & Yes\\
\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
Village & Yes & Yes\\
Year & Yes & Yes\\
\midrule
F (PM 2.5) & 1,369 & 1,168\\
F (PM 2.5 times rice) & 2,107 & 2,201\\
Observations & 1,266,922 & 1,266,922\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). The rice dummy equals one in villages where rice is the most commonly harvested crop. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\begin{figure}
\includegraphics{draft_files/figure-latex/yield8monthly-1} \caption[Effects of pollution on productivity by month]{Effects of pollution on productivity by month}\label{fig:yield8monthly}
\end{figure}

Another way to tease out differential effects on labor productivity, specifically, is to look at the effects of pollution by month. If pollution is working through labor, then we would expect to see larger effects when labor is relatively more important than. The harvest season, the fifth month of the agricultural season in the data, stands out as the most likely candidate for two reasons. First, labor allocation is generally quite on plots during the harvest season. Second, pollution at the very end of the season is unlikely to have large effects on overall land producitivty, as opposed to pollution in prior months.

Figure \ref{yield8monthly} presents the coefficients of the IV regression separately for each month of the agricultural season, with the very beginning of the season on the left side of the graph and the harvest season on the right side of the graph. Interestingly, the coefficient for the fourth month of the season is actually positive. It is hard to know exactly why this is the case, but it could be an interaction between the positive effects of coal pollution on agricultural productivity found by Sanders and Barreca (2022) and the fact that very little labor is used just prior to harvest. However, the overall explanation is beyond the scope of the current paper.

@@

Importantly, since these results come from an IV, both coefficients denote the change in agricultural yield due to an increase in pollution by the same units. In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same amount. These different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}

\hypertarget{quantifying-the-overall-effects-of-pollution-on-agricultural-productivity}{%
\section{Quantifying the overall effects of pollution on agricultural productivity}\label{quantifying-the-overall-effects-of-pollution-on-agricultural-productivity}}

@@

For this reason, I instead turn to data from ICRISAT,\footnote{\url{http://data.icrisat.org/dld}} which has district-level total production data from different years. To track with the timing of the main data used above, I pull total agricultural production from 2001 and 2011 for three of the most common crops in India: rice, wheat, and maize. To simplify the calculation, I assume total area devoted to each crop is unchanged across the decade. I then merge this with the pollution data from Hammer et al. (2020) using a 2011 district-level census shapefile, which has identifiers that match the ICRISAT data. Figure \ref{fig:yielddatamaps} in the appendix shows the distribution of total output for the three crops in 2001. The data covers much, but not all, of the country.

In the output data, I observe values for 2001 and 2011. I can calculate the overall change in log output across the two years, as well as the overall change in pollution levels. I use the most conservative point estimate \emph{without} coal plants in \ref{tab:yield7ivdiffindiff}, which is \(-0.223\). For each district, I calculate the counterfactual for each crop as follows:
\begin{gather} \label{eq:counterfactual} \Delta log(\mathrm{counterfactual})_{d}-0.223\times\Delta log(PM2.5)_{d} = \Delta log(\mathrm{actual})_{d}, \end{gather}

@@

\caption{\label{tab:yieldricestatefe}Agricultural productivity by most common crop}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
  & (1) & (2)\\
\midrule
Particulate matter & -0.811*** & -1.034***\\
(log PM 2.5) & (0.158) & (0.144)\\
PM 2.5 times rice & -1.204*** & -0.966***\\
 & (0.095) & (0.092)\\
Weather (expanded) & Yes & No\\
Weather (expanded, bins) & No & Yes\\
\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
Village & Yes & Yes\\
Year & Yes & Yes\\
\midrule
F (PM 2.5) & 883 & 941\\
F (PM 2.5 times rice) & 1,355 & 1,395\\
Observations & 1,266,922 & 1,266,922\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions estimated agricultural yield (tons per hectare). The rice dummy equals one in villages where rice is the most commonly harvested crop. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\begin{figure}
\includegraphics{draft_files/figure-latex/yielddatamaps-1} \caption[The maps show the district-level yields of each crop in 2001]{The maps show the district-level yields of each crop in 2001.}\label{fig:yielddatamaps}\floatfoot*{Yield data comes from ICRISAT.}
\end{figure}

>>>>

\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper, as well as conference attendees at the 2023 KER International Conference and the University of Reading. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
\author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
\date{2023-12-02}

@@

To identify the effects of air pollution on agricultural productivity, I identify areas of high pollution emission using the location of coal plants. While coal plants may add emissions, I show that they are located in areas with relatively higher levels of pollution, in general. Using these locations as point sources of pollution, I use plausibly exogenous changes in wind direction -- in a similar spirit to Deryugina et al. (2019) -- to measure the effects of pollution on key agricultural outcomes. I first plot the location of coal plants in India by year. Using a village-level shapefile, I identify all villages that are located within 30km of a coal plant at any time during the sample period, which comprises the 2000s to the 2010s. I then calculate the direction from each coal plant to all villages within 30km. I include all areas, regardless of whether the coal plant has opened yet, to allow me to disaggregate the effects of pollution from coal plants with other kinds of pollution later in the paper.

@@

To measure exposure to pollution from point sources around coal plants, I first locate all village centroids located within 30km from a coal plant in a given year. I choose 30km due to previous research on the effects of (air) pollution (Aragón and Rud 2016; Bencko and Symon 1977; Li and Gibson 2014). I then calculate the direction from these point sources to all village centroids within that 30km radius. To define exposure, I pull daily wind direction data from the National Center for Atmospheric Research.\footnote{climatedataguide.ucar.edu/} For each day, I document whether the wind is blowing towards each village centroid\footnote{I define ``towards'' as within five degrees to help take into account that I am using village centroids. The wind data includes both the x and y components of wind speed, which I can use to calculate angles from point sources to village centroids.} and then temporally aggregate this daily data to match the data on productivity. Specifically, agricultural productivity is defined across five months (i.e.~the monsoon season is from June to October) so I count the total days a given village is exposed to wind blowing from \emph{any} coal plant within 30km during those five months.

@@

While I use PM 2.5 as the estimated level of pollution, it is important to clarify that I cannot explicitly say that PM 2.5, per se, is responsible for any effects. Pollution includes a variety of different components, including PM 2.5, PM 10, sulfur dioxide, and nitrogen dioxide, to name but a few. While I use PM 2.5 as a proxy for pollution, it is possible that other components are responsible for the effects. Additionally, the PM 2.5 data is estimated using satellite data and a machine learning algorithm. While this is a common approach in the literature ( see Hammer et al. (2020) for more details), it is important to note that the estimates are not direct measurements of pollution. However, a valid instrumental variable can help correct for any measurement error present in the data (Glewwe 2012; Gillen, Snowberg, and Yariv 2019).

@@

What are the characteristics of areas near the construction of new coal plants? Table \ref{tab:plantresultstable} presents four separate regressions, using state-level fixed effects due to the spatial variation in coal plants. The first two columns use 1991 census data -- downloaded from SHRUG -- and the second two columns use 2001 census data, all at the village level. The first column presents results from regressing a dummy for whether there is a coal plant within the village polygon in 1991 on several census variables. It seems that the exact location of a coal plant does not correlate with overall population of the village, but it does correlate with area; in other words, coal plants are more likely to be located in larger villages but, conditional on area, not in places with more or fewer people.

@@

The second column keeps the 1991 census data but changes the outcome variable to a dummy for whether there \emph{will be} a coal plant constructed within 30km before 2001. This column restricts the sample to only villages that do not already have a coal plant within 30km in 1991. We see that the pattern of coefficients changes from column one to column two; higher populations predict the construction of coal plant in the village over the next ten years. Note that I do not have information on productivity and pollution from 1991.

The last two columns are particularly relevant to the current results. They present similar results to those in the first two columns, but use the 2001 census data and a dummy for the presence of a coal plant in 2001 and 2011 in columns three and four, respectively. Importantly, I include agricultural productivity in the monsoon season of 2002 -- the earliest year available -- as well as average pollution during the monsoon season for 1998 through 2000 as additional predictors. Pollution is a strong predictor for the presence of a coal plant in 2001 as well as for whether a coal plant will be constructed before 2011. While the coefficient is small, it is actually relatively large given that relatively few villages will get a new plant in the next decade. Agricultural producitivty is positively correlated with the presence of a coal plant, but not significantly so.

@@

It is worth taking the time to validate the use of wind direction as a measure of exposure to pollution. Consider the data used for the regressions presented in Table \ref{tab:yield3ivmainfirststage} of the appendix, which is the first stage of the 2SLS procedure. The outcome variable is particulate matter -- specifically, \(\mathrm{PM 2.5}\), which is particulate matter no larger than 2.5 micrometers in diameter -- which comes from Hammer et al. (2020). \(\mathrm{PM 2.5}\) is one of the harmful byproducts from many industrial processes, including power generation. Particulate matter is a major component of the pollution emitted by coal plants, along with sulfur dioxide, different types of nitrogen oxides, and mercury.\footnote{www.epa.gov/airmarkets/power-plants-and-neighboring-communities} This data is available at the monthly level, but I aggregate variables to the season level since that is the level of analysis in the main results.

Both the particulate matter data and the wind data is at the village level, which means the regressions in Table \ref{tab:yield3ivmainfirststage} are at the village level. As such, the geographic fixed effects are village fixed effects and the temporal fixed effects are year fixed effects. Since the exposure variable is at the village level and we follow villages over time, the standard errors are also clustered at the village level. I add more and more flexible specifications for weather, with the most flexible specification in the fourth column. Note that the relationship between wind direction and particulate matter is very similar across all four columns, suggesting that the inclusion of weather does not change the relationship between wind direction and pollution. This is important because it suggests that the use of wind direction as a proxy for pollution exposure is not biased by weather patterns.

@@

Recall, however, the results from above: coal plants were placed in places with higher average agricultural productivity. This means that the simple regressions in Table \ref{tab:yieldtablepm} might actually be biased upwards; although pollution might negatively affect productivity, pollution is higher in more productive areas due to the endogenous placement of pollution sources in the study areas. One possible way to recover the actual effect of pollution on agricultural productivity is with an instrument. If, conditional on the fixed effects and weather variables, wind direction is exogenous and only affects agricultural productivity through pollution, then it is a valid instrument for pollution.

@@

\caption{\label{tab:yieldtableleads}Agricultural productivity and pollution leads}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
\toprule
  & (1) & (2)\\
\midrule
particulate matter (one-year lead) & -0.033 & \\
 & (0.067) & \\
particulate matter (two-year lead) &  & -0.070\\
 &  & (0.052)\\
weather (expanded) & No & No\\
\textbf{fixed effects:} & \textbf{} & \textbf{}\\
village & Yes & Yes\\
year & Yes & Yes\\
\midrule
F & 592 & 783\\
observations & 1,161,265 & 1,055,562\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated (log) agricultural yield (tons per hectare). * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\hypertarget{mechanisms}{%
\subsection{Mechanisms}\label{mechanisms}}

@@

\caption{\label{tab:yieldrice}Agricultural productivity by most common crop}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
  & (1) & (2)\\
\midrule
Particulate matter (log PM 2.5) & -0.287** & -0.394***\\
 & (0.133) & (0.134)\\
PM 2.5 times rice & -0.338*** & -0.205***\\
 & (0.061) & (0.060)\\
Weather (expanded) & Yes & No\\
Weather (expanded, bins) & No & Yes\\
\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
Village & Yes & Yes\\
Year & Yes & Yes\\
\midrule
F (PM 2.5) & 1,369 & 1,168\\
F (PM 2.5 times rice) & 2,107 & 2,201\\
Observations & 1,266,922 & 1,266,922\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated (log) agricultural yield (tons per hectare). The rice dummy equals one in villages where rice is the most commonly harvested crop. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\begin{figure}
\includegraphics{draft_files/figure-latex/yield8monthly-1} \caption[Effects of pollution on productivity by month]{Effects of pollution on productivity by month}\label{fig:yield8monthly}\floatfoot*{Coefficients are from a single regression, where the outcome is (log) agricultural yield and pollution in each of the five months of the season is instrumented with wind in each month. Standard errors are indicated by the bars and are clustered at the village level.}
\end{figure}

Another way to tease out differential effects on labor productivity, specifically, is to look at the effects of pollution by month. If pollution is working through labor, then we would expect to see larger effects when labor is relatively more important. The harvest season, the fifth month of the agricultural season in the data, stands out as the most likely candidate for two reasons. First, labor allocation is generally quite high on plots during the harvest season. Second, pollution at the very end of the season is unlikely to have large effects on overall land producitivty, as opposed to pollution in prior months, since growth is generally complete by the end of the season.

Figure \ref{fig:yield8monthly} presents the coefficients of the IV regression separately for each month of the agricultural season, with the very beginning of the season on the left side of the graph and the harvest season on the right side of the graph. Interestingly, the coefficient for the fourth month of the season is actually positive. It is hard to know exactly why this is the case, but it could be an interaction between the positive effects of coal pollution on agricultural productivity found by Sanders and Barreca (2022) and the fact that very little labor is used just prior to harvest. However, the overall explanation is beyond the scope of the current paper.

@@

Importantly, since these results come from an IV, both coefficients denote the change in agricultural yield due to an increase in pollution by the same units. In other words, this is not identifying differences in overall pollution, per se, but rather differences in the effects of an increase of pollution by the same amount. Since I did not find heterogeneity by exposure, it does not seem that this is driven by non-linear effects of pollution on productivity. Instead, these different effects could be due to, for example, differences in emission patterns for different kinds of plants and factories; coal plants, for example, are responsible for much more mercury than other kinds of polluting industries.\footnote{\url{https://www.epa.gov/sites/default/files/2016-12/documents/nei2014v1_tsd.pdf}}

\hypertarget{overall-effects-of-pollution-on-agricultural-productivity}{%
\section{Overall effects of pollution on agricultural productivity}\label{overall-effects-of-pollution-on-agricultural-productivity}}

@@

For this reason, I instead turn to data from ICRISAT,\footnote{\url{http://data.icrisat.org/dld}} which has district-level total production data from different years. To track with the timing of the main data used above, I pull total agricultural production from 2001 and 2011 for three of the most common crops in India: rice, wheat, and maize.\footnote{Maize might not seem like the first choice for India. However, maize is one of the most common crops grown in other developing countries -- especially in Sub-Saharan Africa -- so I use it here as a benchmark.} To simplify the calculation, I assume total area devoted to each crop is unchanged across the decade. I then merge this with the pollution data from Hammer et al. (2020) using a 2011 district-level census shapefile, which has identifiers that match the ICRISAT data. Figure \ref{fig:yielddatamaps} in the appendix shows the distribution of total output for the three crops in 2001. The data covers much, but not all, of the country.

In the output data, I observe values for 2001 and 2011. I can calculate the overall change in log output across the two years, as well as the overall change in pollution levels. I use the most conservative point estimate \emph{without} coal plants in Table \ref{tab:yield7ivdiffindiff}, which is \(-0.223\). For each district, I calculate the counterfactual for each crop as follows:
\begin{gather} \label{eq:counterfactual} \Delta log(\mathrm{counterfactual})_{d}-0.223\times\Delta log(PM2.5)_{d} = \Delta log(\mathrm{actual})_{d}, \end{gather}

@@

\caption{\label{tab:yieldricestatefe}Agricultural productivity by most common crop}
\centering
\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2cm}}
\toprule
  & (1) & (2)\\
\midrule
Particulate matter & -0.811*** & -1.034***\\
(log PM 2.5) & (0.158) & (0.144)\\
PM 2.5 times rice & -1.204*** & -0.966***\\
 & (0.095) & (0.092)\\
Weather (expanded) & Yes & No\\
Weather (expanded, bins) & No & Yes\\
\textbf{Fixed effects:} & \textbf{} & \textbf{}\\
Village & Yes & Yes\\
Year & Yes & Yes\\
\midrule
F (PM 2.5) & 883 & 941\\
F (PM 2.5 times rice) & 1,355 & 1,395\\
Observations & 1,266,922 & 1,266,922\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item Note: Standard errors are in parentheses and are clustered at the village level. The outcome in all regressions is estimated (log) agricultural yield (tons per hectare). The rice dummy equals one in villages where rice is the most commonly harvested crop. * p<0.10 ** p<0.05 *** p<0.01
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

\begin{figure}
\includegraphics{draft_files/figure-latex/yielddatamaps-1} \caption[District-level yields in 2001]{District-level yields in 2001}\label{fig:yielddatamaps}\floatfoot*{Yield data comes from ICRISAT.}
\end{figure}
